---
title: "04_Prieros_bodymass"
author: "NA"
date: '2023-04-11'
output: html_document
editor_options: 
  chunk_output_type: console
---
Natural variation in body mass of common noctule bats 

Part of the tag mass study, work title:
"Effects of tag mass on the physiology and behaviour of bats: A case study in an insectivorous bat"

This script contains the analysis about the natural variation in body mass of common noctule bats, using a long term data set (10 years) of the colony in Prieros, Brandenburg.

```{r}
# prepare workspace and load packages
pacman::p_load(here, tidyverse, ggplot2, lubridate)

here::here()
path_to_raw <- here("data_raw")
path_to_output <- here("output")

library(ggpubr)
library(extrafont)
```

## load data
```{r}
# data set of colony in Prieros, south Brandenburg
prieros <- read.csv( "data_raw/Prieros_20230301_korr.csv", sep = ";")

# Import via base R if you run into format issues (or look up language format used on the PC and in excel)
```

## Overview
```{r}
str(prieros)

# besides Common noctules data contains observations on P.nat and P.pyg
unique(prieros$species)
prieros %>% count(species)
```

## prepare data
```{r}
# filter for Common noctule bats
nnoc <- prieros %>% filter(species == "Nyctalus noctula")
unique(nnoc$date)
nnoc$date <- dmy(nnoc$date) # change format from "." to "-"
nnoc <- separate(nnoc, date, sep = "-", into = c("year", "month", "day"), remove = FALSE) # create new columns with date split up
str(nnoc)
# set numbers as numeric
nnoc$year <-  as.numeric(nnoc$year)
nnoc$month <- as.numeric(nnoc$month)
nnoc$day <- as.numeric(nnoc$day)

nnoc$forearm <- as.numeric(nnoc$forearm)
nnoc$bodymass <- as.numeric(nnoc$bodymass)

# set male first (as it is normally done in ecology)
nnoc$sex <-  nnoc$sex %>% fct_relevel(c("m", "f"), after = 0) 

# add body condition
nnoc <- nnoc %>% mutate(body.condition = bodymass/forearm)
```

## overview Nnoc data
```{r}
nnoc %>% count(sex)
nnoc %>% count(age)
nnoc %>% count(bodymass)

# exclude data with missing body mass
bm <- nnoc %>% filter(bodymass > 0) # removed 55 NAs
plot(bm$month, bm$bodymass)
summary(bm$bodymass)
plot(bm$month, bm$body.condition)
summary(bm$body.condition)

# filter for adults
bm %>% count(age)
bm.ad <- bm %>% filter(age == "ad.")
plot(bm.ad$month, bm.ad$bodymass)
summary(bm.ad$bodymass)
bm.ad %>% count(month)

############### for intra-individual variation #############
# remove data with missing date
bm.ad <- bm.ad %>% filter(!is.na(date))
# filter for month with lowest body mass (February & March) + highest body mass (August & September)
sub <- bm.ad %>% filter(month == "2" | month == "3" | month == "8" | month == "9")
# alternative filter( ... %in% ...)

str(sub)
sub$sex <- as.character(sub$sex)
sub <- sub %>% select(date, sex, month, bodymass, ring_no, forearm) %>% group_by(ring_no) # keep only useful columns
unique(sub$month) # check

#### population level:
summary(bm.ad$bodymass)
sub %>% filter(month == "2") %>% summary(.)
# bm min in Feb: min 22.0 / max 29.5 / mean 25.6
sub %>% filter(month == "8" | month == "9") %>% summary(.)
# bm max in Aug/Sep: min 24.5 / max 39 / mean 31.1
31.1-25.6 # 5.5
24.5-22   # 2.2
39-29.5   # 9.5

# filter for individuals with data in Feb/Mar + Aug/Sep
sub2 <- sub %>% count(ring_no) %>%  filter(n > 1)
sub3 <- merge(sub, sub2, by.x = "ring_no", by.y = "ring_no")
# write.csv2(sub3, file = "sub3_inclApril.csv")
# no elegant way found in R, manually filtered in excel
# n = 18, maybe include April as well? -> no it's ok, focus on the end of hibernation period

# Individuals with data in winter + summer picked by hand (in excel):
# Feb/Mar & Aug/Sep
sub_list <-  c("A123136", "A123139", "A123160", "A123164", "A123183",
               "A132516", "A132524", "A132538", "A132546", "A132556",
               "A132572", "A132670", "A132680", "A135654", "A135658",
               "A137028", "A137208", "A143604") 

# filter for individuals in the list and create min + max body mass 
sub4 <- sub3 %>% filter(ring_no %in% sub_list) %>% group_by(ring_no) %>% reframe(min = min(bodymass), max = max(bodymass), date = date, sex = sex, forearm = forearm)
# add the gain in body mass
sub4 <- sub4 %>% mutate(diff = max-min)
summary(sub4)
sd(sub4$diff)

# reduce to first row of each ind to remove if individuals measured several times
sub4 <- sub4 %>% distinct(ring_no, .keep_all = TRUE)
sub4 <- sub4 %>% mutate(bc_min = min/forearm, bc_max = max/forearm)
summary(sub4$bc_min); sd(sub4$bc_min) # mean 0.5 + - 0.04 g (min 0.4, max 0.5)
sub4 <- sub4 %>% mutate(rel.diff = diff/min*100)
summary(sub4$rel.diff); sd(sub4$rel.diff) # mean 26.5 + - 11.7 % (min 1.8, max 54,8 %)
summary(sub4$diff); sd(sub4$diff)

# overview male / female
m <- sub4 %>% filter(sex == "m") # n = 6
summary(m$diff); sd(m$diff) # body mass: mean 7 + - 4 g (min 0.5, max 11.5)
summary(m$bc_min); sd(m$bc_min) # bmi: mean 0.5 + - 0.04 g (min 0.4, max 0.5)
summary(m$bc_max); sd(m$bc_max) # bmi: mean 0.6 + - 0.06 g (min 0.5, max 0.6)
f <- sub4 %>% filter(sex == "f") # n = 12
summary(f$diff); sd(f$diff) # bod mass: mean 7 + - 2 g (min 4, max 10.5)
summary(f$bc_min); sd(f$bc_min) # bmi: mean 0.5 + - 0.03 g (min 0.4, max 0.5)
summary(f$bc_max); sd(f$bc_max) # bmi: mean 0.6 + - 0.04 g (min 0.5, max 0.7)
# difference in body.cond
# male
summary(m$bc_max-m$bc_min); sd(m$bc_max-m$bc_min)
# mean diff: 0.13 + - 0.07 / range: 0.01 - 0.22
# female
summary(f$bc_max-f$bc_min); sd(f$bc_max-f$bc_min)
# mean diff: 0.12 + - 0.04 / range: 0.07 - 0.20

# all sex at once
summary(sub4$bc_max-sub4$bc_min); sd(sub4$bc_max-sub4$bc_min)
# mean diff: 0.13 + - 0.05 / range: 0.01 - 0.22

bm.ad %>% count(sex)
summary(bm.ad$date) # Apr 2014 - Feb 2023
# split into male / female
bm.ad.m <- bm.ad %>% filter(sex == "m")
bm.ad.f <- bm.ad %>% filter(sex == "f")
summary(bm.ad.m$bodymass); sd(bm.ad.m$bodymass)
summary(bm.ad.m$body.condition); sd(complete.cases(bm.ad.m$body.condition))
summary(bm.ad.f$bodymass); sd(bm.ad.f$bodymass)
summary(bm.ad.f$body.condition); sd(complete.cases(bm.ad.f$body.condition))
summary(bm.ad$bodymass); sd(bm.ad$bodymass)

# overview repeated measurement data per individual
ind <- bm.ad %>% count(ring_no)
ind %>% filter(n == 1)
# ind 1x 1160 / 2x 319 / 3x 132 / 4x 55 / 5x 35 / 6x 29 / 7x 8 / 8x 6
# 9x 3 / 10x 5 / 11x 1 / 12x 3 / 18x 1 (ring_no: A132692)
unique(ind$n)
barplot(ind$n)

# look at individual measured 18 times
A132692 <- bm.ad %>% filter(ring_no == "A132692")
summary(A132692)

# filter for bats with repeated measurements
ind2 <- ind %>% filter(n > 1)
ring_no <- merge(ind2, bm.ad, by.x = "ring_no", by.y = "ring_no")
unique(ring_no$ring_no)
summary(ind2)
summary(ring_no$bodymass)
bm.ind <- ring_no %>% group_by(ring_no) %>% summarise(mean = mean(bodymass),median = median(bodymass), min = min(bodymass), max = max(bodymass))
bm.ind <- bm.ind %>% mutate(bm.delta = max - min)
ring_no %>% group_by(ring_no)

bm.ad %>% filter(ring_no == "A132692") %>% summary(A132692)
bm.ad %>% filter(ring_no == "A123183") %>% summary(A123183)
bm.ad %>% filter(ring_no == "A132546") %>% summary(A132546)
bm.ad %>% filter(ring_no == "A132670") %>% summary(A132670)

ind3 <- bm.ad %>% filter(ring_no %in% c("A132692", "A123183", "A132546", "A132670"))
summary(ind3)

```

## Pearson Correlation: forearm / bodymass

statistics requirements 
  -assumptions: Continuous data, normally distributed, linearity, 
                no outliers, similar spread across range
                
  Correlation tests
  method = "pearson" [parametric correlation test],
  but method = "kendall" or method = "spearman" [non-parametric correlation test]
  if data do not come from a bivariate normal distribution (more than 2                dimensions, impossible to do some easy plotting)

The Spearman’s rho and Kendall’s tau have the same conditions for use, but Kendall’s tau is generally preferred for smaller samples whereas Spearman’s rho is more widely used.
  
```{r}
# Correlation

hist(ring_no$forearm)
summary(ring_no$forearm)
ring_no %>% filter(forearm > 48) -> ring_no # remove outliers, because expect the three smaller values as measurement errors, e.g. 
hist(ring_no$forearm)
hist(ring_no$bodymass) # normally distributed

hist(ring_no$bodymass, breaks = 20)
qqnorm(ring_no$bodymass); qqline(ring_no$bodymass)
hist(ring_no$forearm, breaks = 20)
qqnorm(ring_no$forearm); qqline(ring_no$forearm)

# or in ggplot
ggqqplot(ring_no$bodymass, ylab = "bodymass")
ggqqplot(ring_no$forearm, ylab = "forearm")

# visualise with scatter
ggscatter(ring_no, x = "bodymass", y = "forearm", add = "reg.line", conf.int = TRUE, cor.coef = TRUE, cor.method = "pearson")
# Shapiro-Wilk to test normality
shapiro.test(ring_no$bodymass) # W = 0.99679, p-value = 0.006805
shapiro.test(ring_no$forearm)  # W = 0.99453, p-value = 6.699e-05

# pearson correlation
cor.test(ring_no$bodymass, ring_no$forearm, method = "pearson", use = "complete.obs")
# if data contains missing values, add: use = "complete.obs"
#cor.test(ring_no$bodymass, ring_no$forearm, method = "kendall", use = "complete.obs")
#cor.test(ring_no$bodymass, ring_no$forearm, method = "spearman", use = "complete.obs")

# interpret results: "soft" positive correlation, cor = 0.3003872, p-value < 2.2e-16
```

--> output: body mass and forearm are positively correlated!
(well, no surprise)


## plot models

Load fonts and set font for ggplots globally
install.packages("extrafont") # see https://github.com/wch/extrafont/

font_import()   # run once
loadfonts()     # loadfonts

to see which are already available: windowsFonts()

```{r}
font_import()
loadfonts(device = "win")
fonts()
theme_set(theme_bw(base_size = 22, base_family = "Bahnschrift"))

#ggplot(bm.ad, aes(x = Fitted, y = Resid)) +
#  geom_point() +
#  facet_wrap(vars(sex)) +
#  theme_bw() +
#  geom_hline(yintercept = 0, col = 'blue')

# plot per sex
names <- c("m" = "Male", "f" = "Female")
count_month_m <- bm.ad %>% filter(sex == "m") %>% count(month)
count_month_m
count_month_f<- bm.ad %>% filter(sex == "f") %>% count(month)
count_month_f


# body mass in g --> used in the article!
theme_set(theme_bw(base_size = 22))
text_x <- c("Feb", "\nMar", "Apr", "\nMay", "Jun", "\nJul", "Aug", "\nSep", "Oct", "\nNov", "Dec") # if used horizontally
text_x <- c("Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec") # if used vertically

ggplot(bm.ad, aes(x = month)) +
  geom_rug(aes(y = bodymass, x = month)) +
  geom_boxplot(aes(y = bodymass, group = month), outlier.shape = 1) +
  #geom_point(aes(y = bodymass), size = 1, shape = 1) +
  facet_wrap(vars(sex), labeller = as_labeller(names)) + # split in female and male
  geom_smooth(aes(y = bodymass, col = sex), method="auto") + # auto chooses "gam"
  scale_x_continuous(breaks = c(2,3,4,5,6,7,8,9,10,11,12), labels = text_x) +
  labs(y = "Body Mass [g]", x = "Month") +
  scale_color_manual(labels = c("male", "female"), values = c("navyblue", "red")) +
  scale_shape_manual(labels = c("male", "female"), values = c(1,4)) +
  guides(colour = "colorbar",size = "none") + # remove legend
  #ylim(18, 42.5) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.3))
  #ggtext::geom_richtext(data = count_month_m, aes(label = n, x = month, y = 42))

# save
#ggsave("plots/bm_prieros.png", plot = last_plot(), width = 900, height = 500, unit = "mm")

# body condition --> maybe supplement?!
ggplot(bm.ad, aes(x = month, y = body.condition)) +
  #geom_point(aes(y = bodymass, col = sex, shape = sex, size = 10)) +
  geom_rug(aes()) +
  geom_boxplot(aes(group = month), outlier.shape = NA) +
  facet_wrap(vars(sex), labeller = as_labeller(names)) +
  geom_smooth(aes(col = sex), method="auto") + # auto chooses "gam"
  scale_x_continuous(breaks = c(2,3,4,5,6,7,8,9,10,11,12)) +
  labs(y = "body condition", x = "month") +
  scale_color_manual(labels = c("male", "female"), values = c("navyblue", "red")) +
  scale_shape_manual(labels = c("male", "female"), values = c(1,4)) +
  guides(colour = "colorbar",size = "none") + # remove legend
  ylim(0.35,0.8)

# female / male in one plot
ggplot(bm.ad, aes(x = month)) +
  #geom_point(aes(y = body.condition, col = sex, shape = sex)) +
  geom_smooth(aes(y = body.condition, col = sex), method="auto") + # auto chooses "gam"
  scale_x_continuous(breaks = c(2,3,4,5,6,7,8,9,10,11,12)) +
  labs(y = "body mass / forearm", x = "month") +
  scale_color_manual(labels = c("female", "male"), values = c("navyblue", "red")) + # wrong?!
  scale_shape_manual(labels = c("female", "male"), values = c(1,4))

# female body mass per month
bm.ad %>% filter(sex == "f") -> female
female %>% filter(month == 9) %>% summary()
# male body mass per month
bm.ad %>% filter(sex == "m") -> male
male %>% filter(month == 12) %>% summary()
# Jul: 30.47, Aug: 31.56, Sep: 30.5, Oct: 30.51, Nov: 30.95, Dec: 29.64 

```

#####plot reg.model + tag mass data in bodymass/forearm

```{r}
theme_set(theme_bw(base_size = 22))

filtered_master_data <- read.csv("output/master_data.csv") # load meta-analysis data
tagged <- filtered_master_data %>% filter(tag.mass > "0")

# rename sex
tagged$sex[ tagged$sex %in% c("m", "Male", "male")] <- "m"
tagged$sex[ tagged$sex %in% c("f", "female")] <- "f"
unique(tagged$sex)

## use this plot to show a well choose data set for meta-analysis
ggplot() + 
  scale_colour_manual(values = c("black", "black", "grey")) +
  geom_point(data = ring_no, aes(bodymass, forearm, fill = "Colony Prieros"), shape = 21, alpha = 0.7, size = 3, fill="darkgrey", colour="black") +
  geom_point(data = tagged, aes(body.mass, forearm, fill = "tagged bats", shape = "tagged bats"), shape = 23, alpha = 0.7, size = 3, fill="#99CCFF", colour="#99CCFF") +
  geom_smooth(data = ring_no, aes(bodymass, forearm), method = "lm", col = "black") +
  #geom_smooth(data = tagged, aes(body.mass, forearm), method = "lm", col = "blue") +
  # not significant!!!
  stat_ellipse(data = ring_no, aes(bodymass, forearm, fill = "Colony Prieros"), level = 0.95, segments = 51, type = "t", size = 1) +
  stat_ellipse(data = tagged, aes(body.mass, forearm, col = "tagged bats"), level = 0.95, segments = 51, type = "t", col = "#99CCFF", size = 1) +
  xlab("body mass [g]") +
  ylab("forearm [mm]")

```
