---
title: "tag_mass_models"
author: "NA"
date: "2024-01-31"
output: html_document
editor_options: 
  chunk_output_type: console
---

Sum up of the tag mass study.

This script contains the analysis and final models of common noctule bats carrying different tag masses, using several studies in a meta-analysis.


## load packages

```{r setup, include=FALSE}

library(readr) # to load data
library(tidyverse)  # contains dplyr (data manipulation) and pipe operator
library(ggplot2) # all kinds of plots
library(ggpubr) # some easy-to-use plots without setting everything as in ggplot2
library(gridExtra) # arrange plots
library(cowplot)  # making effective plot grids
library(ggeffects)  # model predictions
library(broom)  # extracting model summaries
library(effects) # to get the output of the model
library(MuMIn) # for marginal effects
library(lme4) # modelling
library(glmmTMB) # glmm + adjustments
library(DHARMa) # assumption checking
library(report) # summarises facts to report on 
library(here) # to connect to working folder/structure

here::here()
path_to_raw <- here("data_raw")
path_to_output <- here("output")
path_to_plot <- here("plot")

```

## load data

```{r}
filtered_master_data <- read.csv("output/master_data.csv")

# split study date
master_data <- separate(filtered_master_data, study.date, sep = "-", into = c("year", "month", "day"), remove = FALSE)

master_data$year <-  as.numeric(master_data$year)
master_data$month <- as.numeric(master_data$month)
master_data$day <- as.numeric(master_data$day)
# add day of the year
master_data$yday <- yday(master_data$study.date)

# filter to remove urban landscape
#urban <- master_data %>% filter(landscape == "urban") # n = 51
#master_data <- master_data %>% filter(landscape != "urban") # n = 183

```

Data explanations 

str(filtered_master_data)
first.author:          1st author of publication
title:                 publication title
species:               bat species
sex:                   sex of individual bat
age:                   age of individual bat
lat:                   latitude in degree, WGS84 
lon:                   longitude in degree, WGS84
study.location:        summariesed study location (nearby city or village)
study.date:            yyyy-mm-dd
tag.type:              brand, company
tag.mass:              mass of tag + VHF if needed (g)
body.mass:             body mass before tagging (g)
forearm:               forearm length (cm)
body.mass.recapture:   body mass at time of retrieval (g)
date.retrieval:        date of recapture (dd.mm.yyyy)
time.retrieval:        time of recapture (hh:mm)
duration:              time elapsed between tag application and bat retrieval (days)
landscape:             landscape description, most dominant (forest, farmland, urban)
emergence.firstnight:          emergence time during 1st night (min since sunset)
emergence.secondnight:         emergence time during 2nd night (min since sunset)
em.both                        emergence time during 1st night, gaps filled with 2nd night (in case of recording errors or no flight activity in 1 st night)
flight.duration.firstnight:    flight duration (min) of first flight out during first night
flight.duration.secondnight:   flight duration (min) of first flight out during 2nd night
fd.both                        flight duration during 1st night, gaps filled with 2nd night (in case of recording errors or no flight activity in 1 st night)
travelled.distance.firstnight: traveled distance (km) during first flight out during first night
max.distance.roost.firstnight: distance (km) between roost and furthest location away from roost during first night
travelled.distance.secondnight:traveled distance (km) during first flight out during 2nd night
max.distance.roost.secondnight:distance (km) between roost and furthest location away from roost during 2nd night
repository:             repository for flight tracks
ID.bat.tag:                    ID bat/tag
comment                        comments
scheduled_timeinterval_s:      scheduled time interval / ping rate (s)
body.condition                 body mass / forearm length (g/mm)
rel.tag.mass                   tag mass / body mass *100 (%)
type                           relative tag mass category (untagged, < 5 %, 5 to < 10 %, 10 to < 15 %, > 15 %)


## filter data for 1st hypothesis (only data with capture-recapture body mass)

```{r}
# filter body mass data
recap <- master_data %>% filter(body.mass.recap > 0)

# calculate actual weight loss
recap$body.mass.diff <- recap$body.mass.recap - recap$body.mass
summary(recap$body.mass.diff)
# calculate weight loss per day
summary(recap$body.mass.diff/recap$duration)
sd(recap$body.mass.diff/recap$duration)
# calculate relative weight loss
recap$body.mass.rel.diff <- recap$body.mass.diff / recap$body.mass * 100
summary(recap$body.mass.rel.diff)
# calculate relative weight loss per day
summary(recap$body.mass.rel.diff/recap$duration)
sd(recap$body.mass.rel.diff/recap$duration)

# select useful columns
sub_bm <- recap %>% dplyr::select(rel.tag.mass, sex, month, year, study.location, tag.type, body.mass, body.mass.recap, body.condition, duration, landscape, type, body.mass.diff, body.mass.rel.diff)
sub_bm <- sub_bm %>% drop_na()
str(sub_bm)

sub_bm$sex <- as.factor(sub_bm$sex)
sub_bm$landscape <- as.factor(sub_bm$landscape)
sub_bm$type <- as.factor(sub_bm$type)
sub_bm$tag.type <- as.factor(sub_bm$tag.type)
sub_bm$study.location <- as.factor(sub_bm$study.location)
sub_bm$year <- as.factor(sub_bm$year)

# scale values for the model (rel.body.mass, duration, body.condition)
sub_bm$s_rtg <- sub_bm$rel.tag.mass/sd(sub_bm$rel.tag.mass)
```

## filter data for 2nd hypothesis (only data with movement)

```{r}
# filter movement data
move <- master_data
# set as factor
move$sex <- as.factor(move$sex)
move$study.location <- as.factor(move$study.location)
move$landscape <- as.factor(move$landscape)
move$year <- as.factor(move$year) # set year as factor, because we don't expect a relationship between years for the selected values

# emergence time: select useful columns (NA in 1st night filled with 2nd night n=202 / without urban n=178) (first night n=161 / without urban n=137)
sub_em <- move %>% dplyr::select(rel.tag.mass, sex, month, year, study.location, landscape, emergence.firstnight, em.both)
sub_em <- sub_em %>% drop_na(emergence.firstnight)
# flight duration: select useful columns (NA in 1st night filled with 2nd night n=110) (first night n=106)
sub_fd <- move %>% dplyr::select(rel.tag.mass, sex, month, year, landscape, study.location, flight.duration.firstnight, fd.both)
sub_fd <- sub_fd %>% drop_na(fd.both)
```

##  model Recap

```{r}
sub_bm <- as.data.frame(sub_bm)

#Recap <- glmmTMB(body.mass.rel.diff ~ rel.tag.mass + duration + body.condition + landscape + (1|study.location), data = sub_bm, REML = T, family = "gaussian"); summary(Recap)
#simRes <- simulateResiduals(Recap, n = 1000);plot(simRes)

Recap <- lmer(body.mass.rel.diff ~ rel.tag.mass + duration + body.condition + landscape + (landscape|year), data = sub_bm); summary(Recap)
simRes <- simulateResiduals(Recap, n = 1000);plot(simRes)
report(Recap)

# only include terms you are interested in
ggpred_Recap <- ggpredict(Recap, terms = c("landscape")) 
plot(ggpred_Recap)
a <- ggpredict(Recap, terms = c("rel.tag.mass")); plot(a);a
b <- ggpredict(Recap, terms = c("duration")); plot(b);b
c <- ggpredict(Recap, terms = c("body.condition")); plot(c);c

plot(ggpred_Recap) # 1st term = x-achses, 2nd term = colour, 3rd and 4th term = facets

# Plot the predictions with ggplot 
p1 <- ggplot(a) + 
   geom_ribbon(aes(x = x, ymin = predicted - std.error, ymax = predicted + std.error), 
               fill = "lightgrey", alpha = 0.5) +  # error band
   geom_line(aes(x = x, y = predicted), size = 1) + # slope
   geom_point(data = sub_bm, aes(x=rel.tag.mass, y = body.mass.rel.diff), colour = "#666666") +
   labs(x = "Relative Tag Mass [%]", y = "Body Mass Difference [%]", title = " ") + 
   ylim(-25,10)
p1
p2 <- ggplot(b) + 
   geom_ribbon(aes(x = x, ymin = predicted - std.error, ymax = predicted + std.error), 
               fill = "lightgrey", alpha = 0.5) +  # error band
   geom_line(aes(x = x, y = predicted), size = 1) + # slope
   geom_point(data = sub_bm, aes(x=duration, y = body.mass.rel.diff), colour = "#666666") +
   labs(x = "Duration [d]", y = " ", title = " ") + 
   ylim(-25,10) +
  theme(axis.text.y = element_blank(), axis.title.y = element_blank())
p2
p3 <- ggplot(c) +
   geom_ribbon(aes(x = x, ymin = predicted - std.error, ymax = predicted + std.error), 
               fill = "lightgrey", alpha = 0.5) +  # error band
   #geom_point(data = Dat3,                      # adding the raw data (scaled values)
   #           aes(x = rel.tag.mass2, y = body.mass.rel.diff, colour = duration)) +
   geom_line(aes(x = x, y = predicted), size = 1) +  # slope
   geom_point(data = sub_bm, aes(x=body.condition, y = body.mass.rel.diff), colour = "#666666") +
   labs(x = "Body Mass Index [g/mm]", y = " ", title = " ") + 
   ylim(-25,10) +
   coord_cartesian(xlim = c(0.45, 0.74)) +
  theme(axis.text.y = element_blank(), axis.title.y = element_blank())
p3

```


## model Move1 + Move2

```{r}
################### emergence ##########################
Dat2e <- sub_em

# round response variable
Dat2e$em_r <- round(Dat2e$em.both, digits = 0)
summary(Dat2e$em_r)

# model
Move1 <- glmmTMB(em_r+34 ~ rel.tag.mass + sex * month + landscape + (landscape|year), family = Gamma(link = "log"), data = Dat2e); summary(Move1)
simRes <- simulateResiduals(Move1, n = 1000);plot(simRes)
#report(Move1)

# extract predictions for tag mass
(ggpred_Move1 <- ggpredict(Move1, terms = c("rel.tag.mass [2.5:19.5]")))
# subtract the added 34 again
ggpred_Move1$predicted <- ggpred_Move1$predicted-34
ggpred_Move1$conf.low <- ggpred_Move1$conf.low-34
ggpred_Move1$conf.high <- ggpred_Move1$conf.high-34

# plot
e1 <- ggplot(ggpred_Move1) +
   geom_ribbon(aes(x = x, ymin = conf.low, ymax = conf.high), 
               fill = "lightgrey", alpha = 0.5) +  # error band
   geom_line(aes(x = x, y = predicted), size = 1) + # slope
   geom_point(data = Dat2e, aes(x=rel.tag.mass, y = em_r), colour = "#666666") +
   labs(x = NULL, y = "Emergence time \n[min after sunset]", 
        title = " ") +
  theme(axis.text.x = element_blank(), axis.title.x = element_blank())
e1  

# extract predictions for landscape
(ggpred_Move1 <- ggpredict(Move1, terms = c("landscape")))
# subtract the added 34 again
ggpred_Move1$predicted <- ggpred_Move1$predicted-34
ggpred_Move1$conf.low <- ggpred_Move1$conf.low-34
ggpred_Move1$conf.high <- ggpred_Move1$conf.high-34

# plot
land_em <- ggpred_Move1; plot(land_em)
land_em2 <- ggplot(land_em) +
  geom_pointrange(aes(x = x, y = predicted, ymin = conf.low, ymax = conf.high), size = 1, shape = 9) +
  geom_errorbar(aes(x = x, y = predicted, ymin = conf.low, ymax = conf.high), size = 1, width = 0.3) +
  labs(x = NULL, y = "Emergence Time \n[min after sunset]") +
  theme(axis.text.x = element_blank(), axis.title.x = element_blank(), axis.title.y = element_blank())
land_em2

################### flight duration ##########################
Dat2f <- sub_fd

# round response variable (first night because 1st or both are the same data set)
Dat2f$fd_r <- round(Dat2f$fd.both, digits = 0)
summary(Dat2f$fd_r)
# log transform, this time it's possible because we have only positive values for flight duration
Dat2f <- Dat2f %>%
  mutate(log_fd = log(fd_r))

# model
Move2 <- lmer(log_fd ~ rel.tag.mass + sex * month  + landscape + (landscape|year), data = Dat2f); summary(Move2)
simRes <- simulateResiduals(Move2, n = 1000);plot(simRes)
#report(Move2)

# predict tag mass with customized breaks
(ggpred_Move2 <- ggpredict(Move2, terms = c("rel.tag.mass [2.5:19.5]")))
# back transform log
ggpred_Move2$predicted <- exp(ggpred_Move2$predicted)
ggpred_Move2$conf.low <- exp(ggpred_Move2$conf.low)
ggpred_Move2$conf.high <- exp(ggpred_Move2$conf.high)

# plot
f1 <- ggplot(ggpred_Move2) + 
   geom_ribbon(aes(x = x, ymin = conf.low, ymax = conf.high), 
               fill = "lightgrey", alpha = 0.5) +  # error band
   geom_line(aes(x = x, y = predicted), size = 1) + # slope
   geom_point(data = Dat2f, aes(x=rel.tag.mass, y = fd_r), colour = "#666666") +
   xlim(2.5,19.5) +
   labs(x = "Relative Tag Mass [%]", y = "Flight Duration \n[min]", 
        title = " ")
f1

# extract predictions for landscape
(ggpred_Move2 <- ggpredict(Move2, terms = c("landscape")))
# back transform log
ggpred_Move2$predicted <- exp(ggpred_Move2$predicted)
ggpred_Move2$conf.low <- exp(ggpred_Move2$conf.low)
ggpred_Move2$conf.high <- exp(ggpred_Move2$conf.high)

# plot
land_fd <- ggpred_Move2; plot(land_fd)
land_fd2 <- ggplot(land_fd) +
  geom_pointrange(aes(x = x, y = predicted, ymin = conf.low, ymax = conf.high), size = 1, shape = 9) +
  geom_errorbar(aes(x = x, y = predicted, ymin = conf.low, ymax = conf.high), size = 1, width = 0.3, ) +
  labs(x = NULL, y = "Flight Duration \n[min]") +
  scale_x_discrete(labels = c("Farm-\nland", "Forest", "Urban")) +
  theme(axis.title.y = element_blank())
land_fd2
```

### save plots

```{r}
theme_set(theme_classic(base_size = 22))

# body mass
plot_bm <- ggarrange(p1,p2,p3, ncol = 3, nrow = 1, align = "h", label.y = 0, widths = c(1.2,1,1))
plot_bm

# emergence time & flight duration
plot1 <- ggarrange(e1,land_em2, ncol = 2, nrow = 1, align = "h", label.y = 0, widths = c(2,1))
plot1
plot2 <- ggarrange(f1,land_fd2, ncol = 2, nrow = 1, align = "h", label.y = 0, widths = c(2,1))
plot2

# Convert to arrangeGrob
plot_arrangeGrob <- arrangeGrob(plot1,
                                plot2,
                                heights = c(1,1.2),
                                nrow = 2)
# Print the arranged plot
grid.arrange(plot_arrangeGrob)

# resolution for your plots
currres <- 400

### save plot
# adjust ggsave below

```


ggsave(filename = paste("bm2.png", sep = ""), # file name including the ending
       plot = plot_bm, # object that will be saved here
       width = 12, height = 8, # width and height of the plot
       dpi = currres, # resolution
       path = path_to_plot) # path where to save the plot
       
ggsave(filename = paste("em_fd.png", sep = ""), # file name including the ending
       plot = plot_arrangeGrob, # object that will be saved here
       width = 12, height = 8, # width and height of the plot
       dpi = currres, # resolution
       path = path_to_plot) # path where to save the plot
