---
title: "wingspan"
author: "NA"
date: "2023-10-16"
output: html_document
editor_options: 
  chunk_output_type: console
---

Part of the tag mass study, work title:
"Effects of tag mass on the physiology and behaviour of bats: A case study in an insectivorous bat"

This script contains the analysis about the wing span and body mass ratio for the aerodynamic model from measuring bats on the 15-09-2023 in Prieros.
We compare it to the long term data set (10 years) of the colony in Prieros, Brandenburg.

```{r}
# prepare workspace and load packages
pacman::p_load(here, tidyverse, ggplot2, lubridate, ggpubr)

here::here()
path_to_raw <- here("data_raw")
path_to_output <- here("output")

```

## load data
```{r}
# data set of colony in Prieros, south Brandenburg
data <- read.csv("data_raw/Prieros_15-09-23.csv", sep = ";")

# Import via base R if you run into format issues (or look up language format used on the PC and in excel)
```

## Overview
```{r}
# all
hist(data$tip_to_tip_wing_span_cm, breaks = 20)
# adults
ad <- data %>% filter(age == "ad")
hist(ad$tip_to_tip_wing_span_cm)
# subadults
sub <- data %>% filter(age == "sub")
hist(sub$tip_to_tip_wing_span_cm)

# mean + sd per group
group_by(data, age) %>% summarise(count = n(), mean = mean(tip_to_tip_wing_span_cm), sd = sd(tip_to_tip_wing_span_cm) )
# plot both groups
ggboxplot(data, x = "age", y = "tip_to_tip_wing_span_cm", y_lab = "wing span [cm]", xlab = "age")

# wingspan
summary(data$tip_to_tip_wing_span_cm/100); sd(data$tip_to_tip_wing_span_cm/100)
# mean 0.362 m + - 0.11
#aspect ration
summary(data$tip_to_tip_wing_span_cm^2/data$wing_area_cm2); sd(data$tip_to_tip_wing_span_cm^2/data$wing_area_cm2)
# mean 8.04 + - 0.33

# relative tag mass (with 3.8 g tags as used in the wind tunnel experiment)
summary(data$mass_g)
summary(3.8 / data$mass_g * 100); sd(3.8 / data$mass_g * 100)
# mean 13.55 +- 0.83 (range 12.26 - 15.83)
```

## t-test wing span

first test assumptions:
- continuous data
- normally distributed (histogram or Kolmogorov-Smirnov or Shapiro-Wilk test)
- random samples
- enough data (>5 per group, also depending on the research field and expected effect size)
- similar spread between groups

use independent t-test (each individual only measured once)
```{r}
# test normality
# Shapiro-Wilk normality test for adults
with(data, shapiro.test(tip_to_tip_wing_span_cm[age == "ad"]))# p = 0.5
# Shapiro-Wilk normality test for subadults
with(data, shapiro.test(tip_to_tip_wing_span_cm[age == "sub"])) # p = 0.4
# both groups p-values are greater than 0.05 implying that the distribution of the data are not significantly different from the normal distribution

# test variance
res.ftest <- var.test(tip_to_tip_wing_span_cm ~ age, data = data)
res.ftest # p = 0.2
# It’s greater than the significance level alpha = 0.05. In conclusion, there is no significant difference between the variances of the two sets of data.

# t-test
t.test(tip_to_tip_wing_span_cm ~ age, data = data, var.equal = TRUE)
# t = -0.18155, df = 22, p = 0.8576, conf.int[-1.1153973, 0.9358342], est = mean ad: 36.12657 & sub: 36.21635 cm
```

how to read the results:
t-value: test statistics
df = degrees of freedom
p-value
conf.int = confidence interval of the mean 95%
est: sample estimates = mean value of the samples

## t-test wing area
```{r}
# mean + sd per group
group_by(data, age) %>% summarise(count = n(), mean = mean(wing_area_cm2), sd = sd(wing_area_cm2) )
# plot both groups
ggboxplot(data, x = "age", y = "wing_area_cm2", y_lab = "wing area [cm2]", xlab = "age")

# test normality
# Shapiro-Wilk normality test for adults
with(data, shapiro.test(wing_area_cm2[age == "ad"]))# p = 0.7
# Shapiro-Wilk normality test for subadults
with(data, shapiro.test(wing_area_cm2[age == "sub"])) # p = 0.02
hist(sub$wing_area_cm2, breaks = 20)
# both groups p-values are greater than 0.05 implying that the distribution of the data are not significantly different from the normal distribution

# test variance
res.ftest <- var.test(wing_area_cm2 ~ age, data = data)
res.ftest # p = 0.96
# It’s greater than the significance level alpha = 0.05. In conclusion, there is no significant difference between the variances of the two sets of data.

# t-test
t.test(wing_area_cm2 ~ age, data = data, var.equal = TRUE)
# t = -0.3239, df = 22, p-value = 0.7491, conf.int[-8.627624  6.296733], est = ad: 162.3631 & sub: 163.5286
```

results: The p-value of the test is higher than the significance level alpha = 0.05. We can conclude that we couldn't find adults wing span and wing area are  significantly different from subadults measurements. p = 0.86 and p = 0.75.
